---
title: "Figure 5"
author: "Tom Davis"
date: "March 14, 2019"
output: html_document
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "C:\\Users\\thoma\\Desktop\\TNBC_project\\datasets\\TCGA_data")
knitr::opts_chunk$set(fig.width=13.9, fig.height=10.9, message=FALSE, warning = FALSE) 
source("common_functions.R")

```
<br><br><br><br><br><br>



##subplot A-- TMB our dataset
***

data backed up on Xanadu
```{r}

setwd(our_datasets)

snps   <- read_mutation_df("5A_snp.txt", "snps")
indels <- read_mutation_df("5A_indel.txt", "indels")

combined <- full_join(snps, indels)
combined <- add_subtype_our_dataset(combined)
combined <- combined %>% mutate(total=snps+indels)

combined<- combined %>% filter(V2 != "unknown") 
g <- combined  %>% ggplot(aes(x=V4, y=total, fill=V4))+geom_boxplot()+geom_point()+ylab("Somatic Mutation Count ")+xlab("Germline Mutation Status")+stat_compare_means(label.x=1.38, size=STAT_COMPARE_SIZE-5)+coord_cartesian(ylim=c(0,1150))

subplot_A <- my_theme_without_x(g)+theme(axis.text.x=element_text(angle=0))
subplot_A
```

<br><br><br><br><br><br>

##subplot B-- immune expression our dataset
***

expression data was taken from corrected matrix sent by email from ewa on dec 7th 2018
```{r}
setwd(our_datasets)

expression <- read.csv("5B_expression_corrected.csv", stringsAsFactors = FALSE,header = TRUE)


immune_averages <- colMeans(expression[,3:ncol(expression)], na.rm = TRUE) %>% as.data.frame() %>% rownames_to_column()
colnames(immune_averages) <- c("sample", "expression")
immune_with_subtype <- immune_averages %>% mutate(carrier_oneletter=substr(sample,3,3)) %>% filter(carrier_oneletter %in% c("C","N"))
immune_with_subtype <- immune_with_subtype %>% mutate(carrier= ifelse(carrier_oneletter=="C","carrier","noncarrier"))

breast <- immune_with_subtype[(grep("B", immune_with_subtype$sample)),]
ovarian <- immune_with_subtype[(grep("O", immune_with_subtype$sample)),]
breast_plot <- breast %>% ggplot(aes(x=carrier, y=expression, fill=carrier))+geom_boxplot()+geom_point()+xlab("BRCA status")+stat_compare_means( label.x=1,size=7)+ylab("Average Immune Expression (TPM)")

ovarian_plot <- ovarian %>% ggplot(aes(x=carrier, y=expression, fill=carrier))+geom_boxplot()+geom_point()+xlab("BRCA status")+stat_compare_means(label.x=1.5,size=7)+ylab("Average Immune Expression (TPM)")

g <- immune_with_subtype  %>% ggplot(aes(x=carrier, y=expression, fill=carrier))+
  geom_boxplot()+geom_point()+
  labs(x="Germine Mutation Status", y="Immune Expression (TPM)", fill="Germine Mutation Status")+
  stat_compare_means(label.x=.8, size=STAT_COMPARE_SIZE-5)

subplot_B <- my_theme_without_x(g)+theme(axis.text.x=element_text(angle=0))
subplot_B
```

<br><br><br><br><br><br>

##subplot C-- TMB for TCGA TNBC
***
Data from [TCIA](https://tcia.at/neoantigens)
```{r}
setwd(TCGA_datasets)

TMB <- read.table("5C_TCIA_mutations.csv",header=TRUE, stringsAsFactors = FALSE, sep="\t", skip=7)
TMB <- add_subtypes_breast_cbioportal(TMB, 1)
TMB <- TMB %>% filter(!(is.na(TNBCtype_seperated)))
#only want TNBC with subtype information

g <- TMB %>% 
  ggplot(aes(x=TNBCtype_seperated, y=Mutations, fill=carrier))+
  geom_boxplot()+geom_point()+
  geom_hline(yintercept = median(TMB$Mutations, na.rm = TRUE),color="black", linetype = 2, size=HLINE_SIZE)+
  theme_classic()+coord_cartesian(ylim=c(1,170))+
  stat_compare_means(method="wilcox.test", ref.group =".all.",label="..p.signif..", label.y=170, size=STAT_COMPARE_SIZE,  hide.ns = TRUE)+
  labs(fill="Subtype", x="Subtype", y="Somatic Mutation Count")

subplot_C <- my_theme_without_x(g)
subplot_C
```

<br><br><br><br><br><br>

##subplot D -- immune Expression by subtype for TCGA TNBC

expression taken from [Firebrowse](http://firebrowse.org/) 
TODO: this code needs refactoring. Its old and crusty. 

***

```{r}
setwd(TCGA_datasets)

con <- file("5D_immune_breast.txt","r")
first_line <- readLines(con,n=1) %>% strsplit("\t") %>% unlist()
close(con)
expression <- read.table("5D_immune_breast.txt", stringsAsFactors = FALSE, skip=1)
colnames(expression) <- first_line

expression[,2:ncol(expression)] <- apply(expression[,2:ncol(expression)],2, function(x) log2(x+1))
expression_means <- colMeans(expression[,2:ncol(expression)]) %>% as.data.frame()
#log2 transform RSEM normalized counts, and find mean of expression by sample
expression_means <- rownames_to_column(expression_means)
colnames(expression_means) <- c("sample", "expression")

expression_means <- add_subtypes_breast_cbioportal(expression_means,1)

expression_means <- expression_means %>% filter(!is.na(TNBCtype_seperated))

g <- expression_means %>% 
  ggplot(aes(x=TNBCtype_seperated, y=expression, fill=carrier))+
  geom_boxplot()+geom_point(aes(x=TNBCtype_seperated, y=expression))+
  geom_hline(yintercept = mean(expression_means$expression),color="black", linetype = 2, size=HLINE_SIZE)+
  labs(x="Subtype", y="Immune Expression (norm. counts)")+
  theme_classic()+
  stat_compare_means(method="wilcox.test", ref.group =".all.",label="..p.signif..", label.y=8.8, size=STAT_COMPARE_SIZE, hide.ns = TRUE)
subplot_D <- my_theme_without_x(g)
subplot_D
```
<br><br><br><br><br><br>

##subplot E-- LF by subtype for TCGA TNBC

Dataset from supplementary table of [The Immune Landscape of Cancer](https://www.ncbi.nlm.nih.gov/pubmed/29628290)

***

```{r}
setwd(TCGA_datasets)

cibersort <- read.csv("5EH_cibersort_data.csv", stringsAsFactors = FALSE)
cibersort <- cibersort %>% filter(TCGA.Study=="BRCA") %>% dplyr::select(TCGA.Participant.Barcode
,Leukocyte.Fraction)

cibersort <- add_subtypes_breast_cbioportal(cibersort,1)



subplot_E <- cibersort %>% filter(!is.na(TNBCtype_seperated)) %>% 
  ggplot(aes(x=TNBCtype_seperated,y=Leukocyte.Fraction, fill=carrier))+
  geom_boxplot()+geom_point()+
  geom_hline(yintercept = mean(cibersort$Leukocyte.Fraction, na.rm = TRUE),color="black", linetype = 2, size=HLINE_SIZE)+
    labs(x="Subtype", y="Leukocyte Fraction")+
  theme_classic()+
  stat_compare_means(method="wilcox.test", ref.group =".all.",label="..p.signif..", size=STAT_COMPARE_SIZE,  hide.ns = TRUE)


subplot_E <- my_theme_without_x(subplot_E)
subplot_E
```

<br><br><br><br><br><br>

##subplot F-- TMB ovarian TCGA OV 

Data from [TCIA](https://tcia.at/neoantigens)
***

```{r}
setwd(TCGA_datasets)

OV <- read.table("5F_TCIA_mutations_OV.csv", sep="\t", skip=7, header=TRUE, stringsAsFactors = FALSE)
OV <- add_subtypes_ovarian(OV, 1) 

OV_TMB <- OV %>% ggplot(aes(x=SUBTYPE, y=Mutations, fill=carrier))+
  geom_boxplot()+geom_point()+
  geom_hline(yintercept = median(TMB$Mutations, na.rm = TRUE),color="black", linetype = 2, size=HLINE_SIZE)+
  labs(fill="Subtype", x="Subtype", y="Somatic Mutation Count")+
  theme_classic()+
  stat_compare_means(method="wilcox.test", ref.group =".all.",label="..p.signif..", size=STAT_COMPARE_SIZE,  hide.ns = TRUE)
subplot_F <- my_theme_without_x(OV_TMB)
subplot_F
```

<br><br><br><br><br><br>

##subplot G-- immune expression by subtype for TCGA OV 
Data from [firebrowse](http://firebrowse.org/) downloaded from [xena](https://xena.ucsc.edu/)

***

```{r}
setwd(TCGA_datasets)
con <- file("5G_immune_ovarian.tsv","r")
first_line <- readLines(con,n=1) %>% strsplit("\t") %>% unlist()
close(con)
ovarian <- read.table("5G_immune_ovarian.tsv", sep="\t", skip=1, stringsAsFactors = FALSE)
colnames(ovarian) <- first_line
ovarian<- ovarian %>% dplyr::select(-contains("unknown"))
ovarian_rowmeans <- data.frame(ID=ovarian[,1], expression=rowMeans(ovarian[,2:ncol(ovarian)]), stringsAsFactors = FALSE)

ovarian_rowmeans<- ovarian_rowmeans %>% filter(substr(ID, 14,15)=="01") %>%        mutate(ID=substr(ID, 1,12))

ovarian_rowmeans <- add_subtypes_ovarian(ovarian_rowmeans, 1) %>% na.omit()

ovarian_rowmeans_plot <- ovarian_rowmeans %>% 
  ggplot(aes(x=SUBTYPE,y=expression, fill=carrier))+
  geom_boxplot()+geom_point()+
  geom_hline(yintercept = mean(ovarian_rowmeans$expression),color="black", linetype = 2, size=HLINE_SIZE)+
  theme_classic()+
  labs(x="Subtype", y="Immune Expression (norm. counts)")+
  stat_compare_means(method="wilcox.test", ref.group =".all.",label="..p.signif..", label.y=8.5, size=STAT_COMPARE_SIZE,  hide.ns = TRUE)
subplot_G<- my_theme_without_x(ovarian_rowmeans_plot)
subplot_G

```

<br><br><br><br><br><br>

##subplot H-- LF by subtype for TCGA OV

Dataset from supplementary table of [The Immune Landscape of Cancer](https://www.ncbi.nlm.nih.gov/pubmed/29628290)

***

```{r}
setwd(TCGA_datasets)
cibersort <- read.csv("5EH_cibersort_data.csv", stringsAsFactors = FALSE)
ovarian <- cibersort %>% filter(TCGA.Study=="OV")
ovarian <- add_subtypes_ovarian(ovarian, 1)

ov_LF <-  ovarian %>% dplyr::select(ID, carrier, Leukocyte.Fraction, SUBTYPE) %>% na.omit()

ov_LF_plot <- ov_LF %>% 
  ggplot(aes(x=SUBTYPE,y=Leukocyte.Fraction, fill=carrier))+
  geom_boxplot()+geom_point()+geom_hline(yintercept = mean(ovarian$Leukocyte.Fraction, na.rm = TRUE),color="black", linetype = 2,size=HLINE_SIZE)+
  stat_compare_means(method="wilcox.test", ref.group =".all.",label="..p.signif..", size=STAT_COMPARE_SIZE, hide.ns = TRUE)+
  theme_classic()+
  labs(x="Subtype", y="Leukocyte Fraction")
subplot_H <- my_theme_without_x(ov_LF_plot)
subplot_H
```

If you want as separate figures
```{r}
setwd("C:\\Users\\thoma\\Desktop\\TNBC_project\\figures")

graphs <- list(subplot_A, subplot_B, subplot_C, subplot_D, subplot_E, subplot_F, subplot_G, subplot_H)

save_plot_list(graphs)

 ab_legend <- get_legend(subplot_B+ theme(legend.position="right", legend.key.size = unit(2, "cm"), legend.title = element_text(face="bold",size=FONTSIZE-20, color="black"), legend.text = element_text(face="bold",size=FONTSIZE-20, color="black")))
ggsave(ab_legend, filename = "legend_file.png")
getwd()


```

If you want as cowplot: 

```{r, warning=FALSE}
# setwd(FIGURE_IMAGE_WD)

# 
# ab <- cowplot::plot_grid(subplot_A, subplot_B, ab_legend, labels=c("A","B"), nrow=1, label_size=LABELSIZE, scale=COWPLOT_SCALE)
# 
# 
# cde <- 
#   cowplot::plot_grid(subplot_C, subplot_D , subplot_E, labels=c("C","D", "E"), nrow=1, ncol=3, label_size=LABELSIZE,  scale=COWPLOT_SCALE)
# fdg <- cowplot::plot_grid(subplot_F, subplot_G , subplot_H, labels=c("F","G", "H"), nrow=1, ncol=3, label_size=LABELSIZE,  scale=COWPLOT_SCALE)
# combined <- cowplot::plot_grid(ab, cde, fdg, nrow=3)
# cowplot::save_plot("fig5.png", combined, base_height=46, base_aspect_ratio = 1.2, limitsize=FALSE)
```


